---
title: "Who does more pushups"
author: "your name"
date: "March 8"
output: html_document
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

Several years ago some MAT 111 students who were on the football team conducted a study of their team-mates.  The results are contained in the data frame `pushups` in the `tigerstats` package.

```{r eval=FALSE}
data(pushups)
View(pushups)
help(pushups)
```

The data frame created from the MAT 111 study contains observations of thirty football players based on three variables: weight of the player, number of pushups performed in a two minute period, and the position played (either line or skill). 

Line players are on the offensive or defensive line.  They need to be a certain kind of speed, but they must also be quite big and strong.  Skill players do a lot of running and catching:  they are the quarterback, running backs, receivers, etc.

The goal of this report is to examine and compare the number of pushups a player can perform based on their position in order to answer the following research question:

  >Who does more pushups, on average: a line player or a skill player?


# Methods
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

To examine the MAT 111 football pushup study, I will utilize a box plot as a graphical method and a variable analysis as a numerical method. The variable analysis will include comparing the factor variables of football player positions with the numerical factors of amount of pushups performed in a two minute period. In addition to these methods, I will form hypotheses about the relationship between the variables and use a t-test to determine which hypothesis to accept. 

# Results
<!--You have both graphical and numerical methods, so you will have both graphical and numerical results to consider.!-->


## Graphical Results
<!--Your graph should have a good title, and axes should be labled as needed.  Discuss your graphical results, showing how they answer the research question.  Use specific features of the graph in your discussion.!-->
Looking at the boxplot below, we can determine whether the amount of pushups performed by each type of player appears to be the same.

```{r}
bwplot(pushups~position,data=pushups,
main="Who Does More Pushups?",
xlab="Player Position",
ylab="Pushups")
```

The boxplot of each player position are located at different points along the y-axis, indicating that the number of pushups performed by each position is different. The upper and lower whiskers on the plot reveal that the range of pushups is smaller for line players than for skill players. Line players appear to perform anywhere from 5 to 35 pushups; the skill players appear to perform anywhere from 20 to 90 pushups, with one player performing over 100 pushups. 

More importantly, the dots inside the boxes show the median number of pushups performed. For line players, the median appears to be about 18 pushups; whereas the median for skill players is about 50 pushups. These numbers reveal that on average, the skill players perform more pushups. 

## Numerical Results
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

It is important to verify the indicated results of our boxplot by looking at exact numbers.

```{r}
favstats(pushups~position,data=pushups)
```

The data above is in line with our interpretation of the box plot. The average number of pushups performed by line players is 17.13, and the average number of pushups perfomred by skill players is 49.73. These significantly different numbers further our conclusion that skill players do, in fact, perform more pushups.


# Inferential Results
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
<!--State your overall conclusion.  This is also a good place to mention other Research Questions that came to mind as you were carrying out this study.!-->

