---
title: "Who does more pushups"
author: "your name"
date: "March 8"
output: html_document
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

Several years ago some MAT 111 students who were on the football team conducted a study of their team-mates.  The results are contained in the data frame `pushups` in the `tigerstats` package.

```{r eval=FALSE}
data(pushups)
View(pushups)
help(pushups)
```

The data frame created from the MAT 111 study contains observations of thirty football players based on three variables: weight of the player, number of pushups performed in a two minute period, and the position played (either line or skill). 

Line players are on the offensive or defensive line.  They need to be a certain kind of speed, but they must also be quite big and strong.  Skill players do a lot of running and catching:  they are the quarterback, running backs, receivers, etc.

The goal of this report is to examine and compare the number of pushups a player can perform based on their position in order to answer the following research question:

  >Who does more pushups, on average: a line player or a skill player?


# Methods
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


# Results
<!--You have both graphical and numerical methods, so you will have both graphical and numerical results to consider.!-->


## Graphical Results
<!--Your graph should have a good title, and axes should be labled as needed.  Discuss your graphical results, showing how they answer the research question.  Use specific features of the graph in your discussion.!-->
Looking at the boxplot below, we can determine whether the amount of pushups performed by each type of player appears to be the same.

```{r}
bwplot(pushups~position,data=pushups,
main="Who Does More Pushups?",
xlab="Player Position",
ylab="Pushups")
```

The boxplot of each player position are located at different points along the y-axis, indicating that the number of pushups performed by each position is different. The upper and lower whiskers on the plot reveal that the range of pushups is smaller for line players than for skill players. Line players appear to perform anywhere from 5 to 35 pushups; the skill players appear to perform anywhere from 20 to 90 pushups, with one player performing over 100 pushups. 

More importantly, the dots inside the boxes show the median number of pushups performed. For line players, the median appears to be about 18 pushups; whereas the median for skill players is about 50 pushups. These numbers reveal that on average, the skill players perform more pushups. 

## Numerical Results
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

It is important to verify the indicated results of our boxplot by looking at exact numbers.

```{r}
favstats(pushups~position,data=pushups)
```

The data above is in line with our interpretation of the box plot. The average number of pushups performed by line players is 17.13, and the average number of pushups perfomred by skill players is 49.73. These significantly different numbers further our conclusion that skill players do, in fact, perform more pushups.


# Inferential Results
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

The use of a t-test will allow us to compare the means we have derived from our data set. We want to determine if there is a relationship between the amount of pushups a player can perform and the position they play. First, we must state our null and alternative hypotheses.

$H_0$:  There is no relationship between the amount of pushups a player can perform and the position they play. 

$H_a$: There *is* a relationship between the amount of pushups a player can perform and the position they play. 

```{r}
t.test(pushups~position,data=pushups)
```

The difference in means calculated in this test is not equal to zero, meaning there is a difference between the two groups of players and the pushups they can do. The p-value from this t-test is 0.000009421. If the null hypthesis is indeed true, then the results we just obtained would happen about far less than one percent of the time. Such a small number means we must reject the null hypothesis and conclude that the number of pushups a player can perform is related to the position they play.


# Discussion and Conclusion
<!--State your overall conclusion.  This is also a good place to mention other Research Questions that came to mind as you were carrying out this study.!-->

After analyzing the data both graphically and numerically, we can conclude that skill players do, on average, perform more pushups than line players. The box plot allowed for a visual representation of the differences in the number of pushups, and the exact numbers of pushups performed and the means confirmed those differences.  

Futhermore, the p-value obtained from the t-test, led us to reject our null hypothesis and instead favor the alternative hypothesis that there **is** a relationship between the number of pushups a football player can perform and the position they play. 

While carrying out this study, we did not examine the variable of football players' weight as recorded by the MAT 111 students. It would be interesting to conduct further research to determine if the weight of a player has any bearing on the amount of pushups they can perform. For example, even though we concluded that the average skill player can perform more pushups, we could research which skill player can perform more: a heavier weight or a lighter weight skill player.
